---
title: "Preliminaries"
author: "Albert Rapp"
date: June 19, 2024
format:
  html:
    code-link: true
    code-fold: true
    self-contained: true
    embed-resources: true
    toc: true
    theme: Zephyr
    number-sections: false
execute: 
  warning: false
  message: false
  echo: false
editor_options: 
  chunk_output_type: console
---`
---

# Preliminaries

In this session, Albert provides an overview of the main functions and shortcuts to be used through the course.

## Lesson 1

Running the code

```{r running_code}
#| echo: true

library(tidyverse)
library(janitor)
library(skimr)
library(palmerpenguins)

penguins |> 
  skimr::skim()
modeldata::ames |> 
  janitor::clean_names() |> 
  skimr::skim(lot_frontage, lot_area)



```

## Displaying output

Enable to display the output in the **Console(Viewer screen)** instead of the **Inline**

## Referring to documentation on the fly

Highlight the command or function interested in, then press `fn` (function key) along with `F1` key.

## Install Instructions

The install instructions are available at this [website](file:///Users/albertofcabrera/Documents/R-books/3MW(Albert%20Rapp)/r_data_cleanning_master_class/Lbcp8OXMTZ6Y3iPZ9Ui4_downloads/00_getting_started.html).

```{r packages needed}

pkgs <- c('tidyverse', 'palmerpenguins', 'here', 'janitor', 'modeldata', 'skimr', 'quarto', 'sessioninfo')


install.packages(pkgs)


```
